Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-03 paper/paper.tex
Add text in conclusions

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
\end{center}
\caption{Cross-section of the wavefield used as a reference sample for compression and decompression}
\label{fig:uncompressed}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/errors.pdf}
\end{center}
\caption{Cross-section of the field that shows errors introduced during compression and decompression using the fixed-tolerance mode}
\label{fig:decompressed_error}
\end{figure}

@@

\section{Results}
Figure \ref{fig:varying_memory} shows the speedups predicted by the
performance model for varying amount of available memory. A big
shortcoming that was realised in ZFP was the lack of parallelization
in the decompression functions. The ZFP manual talks about the
difficulty in implementing a parallel mode decompressor. To estimate
the benefit from parallel decompression, we scaled the serial
decompression time by the same factor that we saw between the serial
and parallel compressors on a given platform. We called this scaled
decompression time, ``theoretical decompression time'' and show this
on the plot with a dotted line. The
right-most region of this plot has been the focus of many studies
mentioned earlier in this paper. The extension to the regions with
lower memory is new here and we see that the maximum benefit is gained
in the most memory-constrained scenarios (region 1), while there is a
region where Revolve-based recomputation alone would be more
beneficial (region 2). As expected, compression alone (with no
Revolve) dominates region 3. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-memory.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  memory. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The dotted curve represents the hypothetical speedup
if ZFP supported parallel decompression. The two vertical lines at 20GB and 52GB clearly
divide the plot into 3 regions. The left-most region, with less than
20GB memory sees the maximum benefit from compression. The middle
region, with memory between 20GB and 52GB, is better off using only
Revolve and no compression. The right-most region, with memory greater
than 52GB represents the scenario where the entire history of the
wavefield can fit in memory post-compression. In this case compression
eliminates the use of Revolve - hence the speedup. This is the region
that has been well-studied in past work.}
\label{fig:varying_memory}
\end{figure}

@@

Figure \ref{fig:varying_nt} shows the speedups predicted for differing
number of timesteps. It is interesting to note that there is a
consistent increase in the speedup as the number of timesteps goes
up. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}
\section{Conclusions and Future work}

What are the problems?

Future work:
\begin{itemize}
\item Acceptable error bounds
\item A posteriori and a priori metrics
\item Relationship between compression error and field error, propagation of errors in forward simulation
\item Relationship between compression error and gradient error in cross-correlation
\item What else?
\end{itemize}

More than 3D fields (use time correlation)

Parallel decompression

Disk checkpointing

\section*{Acknowledgment}
This work was funded by the Intel Parallel Computing Centre at
Imperial College London and EPSRC EP/R029423/1 . 
This work was supported by the U.S. Department of Energy, Office of Science,
Office of Advanced Scientific Computing Research, Applied Mathematics and
Computer Science programs under contract number DE-AC02-06CH11357.
We would also like to acknowledge the support from the SINBAD II project and
the member organizations of the SINBAD Consortium.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
\end{center}
\caption{Cross-section of the wavefield used as a reference sample for
  compression and decompression. This field was formed after a Ricker
  wavelet source was placed at TODO and the wave propagated for 2500
  timesteps. This is a vertical cross-section of a 3D field, taken at
  the source location}
\label{fig:uncompressed}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/errors.pdf}
\end{center}
\caption{Cross-section of the field that shows errors introduced
  during compression and decompression using the fixed-tolerance
  mode. It is interesting to note that the errors are more or less
  evenly distributed across the domain with only slight variations
  corresponding to the wave amplitude (from the field plot in figure
  \ref{uncompressed}. A small block-like structure characteristic of
  ZFP can be seen.}
\label{fig:decompressed_error}
\end{figure}

@@

\section{Results}
We notice that the combination of compression and checkpointing
outperforms pure checkpointing in some scenarios, and that wherever
possible, compression alone outperforms checkpointing. This is made
evident in figure \ref{fig:varying_memory} which shows a clear split
in three different regions. The right-most region of this plot has
been the focus of many studies mentioned earlier in this paper. The
extension to the regions with lower memory is new here and we see that
the maximum benefit is gained in the most memory-constrained scenarios
(region 1), while there is a region where Revolve-based recomputation
alone would be more beneficial (region 2). As expected, compression
alone (with no Revolve) dominates region 3.
Figure \ref{fig:varying_memory} shows the speedups predicted by the
performance model for varying amount of available memory. A big
shortcoming that was realised in ZFP was the lack of parallelization
in the decompression functions. The ZFP manual talks about the
difficulty in implementing a parallel mode decompressor. To estimate
the benefit from parallel decompression, we scaled the serial
decompression time by the same factor that we saw between the serial
and parallel compressors on a given platform. We called this scaled
decompression time, ``theoretical decompression time'' and show this
on the plot with a dotted line. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-memory.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  memory. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The dotted curve represents the hypothetical speedup
if ZFP supported parallel decompression. The two vertical lines at 20GB and 52GB clearly
divide the plot into 3 regions. The left-most region, with less than
20GB memory sees the maximum benefit from compression. The middle
region, with memory between 20GB and 52GB, is better off using only
Revolve and no compression. The right-most region, with memory greater
than 52GB represents the scenario where the entire history of the
wavefield can fit in memory post-compression. In this case compression
eliminates the use of Revolve - hence the speedup. This is the region
that has been well-studied in past work.}
\label{fig:varying_memory}
\end{figure}

@@

Figure \ref{fig:varying_nt} shows the speedups predicted for differing
number of timesteps. It is interesting to note that there is a
consistent increase in the speedup as the number of timesteps goes
up. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}
\section{Conclusions and Future work}
What are the problems?
The performance we developed might be useful in judging whether compression
would be helpful in a particular scenario without actually building a
complete implementation first. We have tried to keep it application
and platform agnostic in order to maximize its applicability. However
there are a few limitations in this model. Firstly, we assumed that
the compression performance we saw for a single checkpoint is
representative of any checkpoint throughout the simulation. This
assumption might have been completely wrong if made about a checkpoint
in the first few time steps since we consistently saw upto 800x
compression in those cases - since they are mostly zeros. In order for
this assumption to be realistic, we use a checkpoint close to the end
of the simulation so that most of the field is perturbed and hence is
close to the worst case in terms of compression results.

In comparing pure compression with pure checkpointing, we assume that every
checkpoint would be compressed and decompressed. However, in scenarios
where the available memory is only slightly less than the required
memory, an implementation that compresses only a subset of the
checkpoints might outperform the expectations of our model.

Future work:

This work is focussed on checkpointing in memory alone. In a future
work we would like to extend this performance model to include
checkpointing to disk.

The ZFP algorithm that the majority of this paper was based on, only
supports fields with up to 3 dimensions. Since our field is evolving
in time, we would expect some correlations in the time dimensions that
are currently not being exploited in the compression.

As can be seen in figure \ref{fig:varying_memory}, the lack of
parallel decompression has a significant impact on the overall time to
solution. This strategy would become much more beneficial once we
start seeing faster decompression speeds.

In this work, we have ignored the error that is induced by the lossy
compression - assuming that a tolerance of $10^{-4}$ is acceptable for
the application. This assumption needs to be verified in a formal
analysis of how the error in the field propagates through the
simulation, and how it finally affects the cross-correlated gradient. 
The development of a priori and a posteriori error metrics would
greatly help the further use of lossy compression in similar
problems. 
\begin{itemize}
\item Acceptable error bounds
\item A posteriori and a priori metrics
\item Relationship between compression error and field error, propagation of errors in forward simulation
\item Relationship between compression error and gradient error in cross-correlation
\item What else?
\end{itemize}

\section*{Acknowledgment}
This work was funded by the Intel Parallel Computing Centre at
Imperial College London and EPSRC EP/R029423/1. 
This work was supported by the U.S. Department of Energy, Office of Science,
Office of Advanced Scientific Computing Research, Applied Mathematics and
Computer Science programs under contract number DE-AC02-06CH11357.
We would also like to acknowledge the support from the SINBAD II project and
the member organizations of the SINBAD Consortium.
